---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r echo=TRUE}
library(tidyverse)
library(dplyr)
library(xtable)
data <- read.csv(unz("activity.zip","activity.csv"))
data <- mutate(data, date = ymd(date))
data <- group_by(data, by = day(date))

```

## Making histogram

```{r}
library(ggplot2)
graph <- ggplot(data,aes(x = date, weight = steps))
graph + geom_histogram(binwidth = 1) + labs(title = "Steps taken per day", y = "Steps",
        x = "") + theme(plot.title = element_text(hjust = 0.5))
```

## What is mean total number of steps taken per day?

```{r, results= "asis"}
library(xtable)
mean <- summarise(data, mean = mean(steps, na.rm = TRUE))
mean <- mutate(mean, median = summarise(data, median = median(steps, na.rm = TRUE))$median)
mean <- xtable(mean[,-1])
print(mean,type = "html")
```

## What is the average daily activity pattern?

```{r}
data <- mutate(data, interval = minutes(interval))
new_data <- group_by(data, interval)
interval_mean <- summarise(new_data, steps = mean(steps, na.rm = TRUE))
line_plot <- ggplot(interval_mean, aes(x = minute(interval), y = steps))
line_plot + geom_line()  + labs(x = "5 Minutes Intervel")

```

#Max no of steps taken on average across all days

```{r}
max <- max(interval_mean$steps)
interval_mean[which(interval_mean$steps == max),]
```

##Calculating Missing values

```{r}
mean(is.na(data$steps)) *100

```
There are 13.12% missing values in the data

## Imputing missing values

```{r}

no_nas <- new_data %>% mutate(steps = replace_na(steps, as.integer(round(mean(steps, na.rm = TRUE)))))
```

In the above chunk I have calculated the mean for the 5 minute interval and replaced the nas with it.

```{r, results= "asis"}

average <- no_nas %>% group_by(day(date))%>% summarise(mean = mean(steps))
average <- no_nas %>% group_by(day(date)) %>% summarise(median = median(steps)) %>% mutate(average,median = median)
average <- xtable(average[,-1])
print(average, type = "html")

```

```{r}
graph <- ggplot(no_nas,aes(x = date, weight = steps))
graph + geom_histogram(binwidth = 1) + labs(title = "Steps taken per day", y = "Steps",
        x = "") + theme(plot.title = element_text(hjust = 0.5))

```

## Are there differences in activity patterns between weekdays and weekends?

```{r}

by_week <- mutate(no_nas,
        day_type = if_else(weekdays(date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday"),
        day_type = factor(day_type, levels = c("Weekday", "Weekend")))
by_week <- group_by(by_week, day_type, interval) %>% summarise(steps = mean(steps), .groups = "drop_last")
line_grap <- ggplot(by_week, aes(x = as.numeric(interval, "minutes"), y = steps, group = 1))
line_grap + geom_line() + facet_grid(by_week$day_type~.)

```
